---
title: "Osteosarcoma - lung epithelial interactions"
author: "Ryan Roberts"
date: March 2022
output: knitrBootstrap::bootstrap_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
    cache = TRUE,
    bootstrap.show.code = FALSE,
    bootstrap.show.output = FALSE,
    bootstrap.show.message = FALSE)
```

# Osteosarcoma - lung epithelial interactions

## Goals:

Generate figures for the tumor-epithelial paracrine signaling, self-seeding mechanisms paper.
This particular file is being used to consolidate the raw data into organized R data objects.

## Samples:

placeholder

## Approach

placeholder

# Workflow

## Load libraries
```{r libraries, cache=FALSE, eval=TRUE}
library(tidyverse)
library(pzfx)
```

## Helper functions

placeholder

## Get list of data sources
```{r}
data_sources <- readxl::read_excel("D:/OneDrive - Nationwide Children's Hospital/Papers/Active Manuscripts/Self seeding and IL6-IL1 paracrine loop/Source Data/Locations and Names of Data.xlsx")
```

## Cytokine production in OS-epithelial co-cultures with inhibitors
Taken from file "PAPER FIGURES-CXCL8-IL6 data.pzfx"
```{r}
cytokine_cocx <- list()

cytokine_cocx$HBEC$OS-17$IL6 <- tribble(

)
```





## Figure 4: Import transwell data and create data object
Save data objects with metadata to data/
Compare to media only

```{r plotCytokines, eval=TRUE}
paper_path <- "../Source Data/"

migration <- list()

migration$cytokine <- list()
migration$cond_media <- list()

migration$cytokine$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 Migration", "metastatic", "to_cytokine",
    "143B", "143B Migration", "metastatic", "to_cytokine",
    "K7M2", "K7M2 Migration", "metastatic", "to_cytokine",
    "OS-25", "OS-25 Migration", "non-metastatic", "to_cytokine",
    "OSCA8", "OSCA8 Migration", "", "to_cytokine")

migration$cond_media$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 migration assay (6.14.18) to inhibitors",
        "metastatic", "to_inhibitors",
    "143B", "143B migration (2018-06-01) to inhibitors",
        "metastatic", "to_inhibitors",
    "K7M2", "K7M2 migration (2018-06-26) to inhibitors",
        "metastatic", "to_inhibitors",
    "SaOS2", "SaOS2 migration (2018-07-11) to inhibitors",
        "non-metastatic", "to_inhibitors",
    "MG63", "MG-63 migration (#1 2018-05-25) to inhibitors",
        "non-metastatic", "to_inhibitors")

migration$cytokine$data <- lapply(
    migration$cytokine$source$table_name, 
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(
            description = x,
            met_potential = migration$cytokine$source$met_potential[
                migration$cytokine$source$table_name == x]) %>%
        filter(condition != "CXCL12" & condition != "OSM") %>%
        mutate(condition =
            str_replace(condition, "^Media$", "Media") %>%
            str_replace("1% FBS", "FBS")) %>%
        mutate(condition = factor(condition, 
            levels = c("Media", "FBS", "IL6", "IL8")))
    })
names(migration$cytokine$data) <- migration$cytokine$source$cell_line

migration$cond_media$data <- lapply(
    migration$cond_media$source$table_name,
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(description = x,
            met_potential = migration$cond_media$source$met_potential[
                migration$cond_media$source$table_name == x])
        ))
)
        if (data_type == "to_cytokine") {
            dataset <- dataset %>%
            
    } else if (data_type == "to_inhibitors") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("Cond Media", "Cond. media (C.M.)") %>%
                       str_replace("CM \\+ sc144", "C.M. + SC144") %>%
                       str_replace("CM \\+ ladarixin", "C.M. + DF2156A") %>%
                       str_replace("CM \\+ both", "C.M. + both")) %>%
            mutate(condition = factor(condition, levels = c('Media only',
                                                            '1% FBS',
                                                            'Cond. media (C.M.)',
                                                            'C.M. + SC144',
                                                            'C.M. + DF2156A',
                                                            'C.M. + both')))
    } else if (data_type == "to_conditioned_media") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("1% FBS", "Media + FBS") %>%
                       str_replace("Cond Media", "Cond. media")) %>%
            mutate(condition = factor(condition, levels = c("Media only",
                                                            "Media + FBS",
                                                            "Cond. media")))
    }

    cmp_group <- if_else(data_type == "to_conditioned_media",
                         true = "Media . FBS",
                         false = "Media only")
    aov_test <- aov(cells ~ condition, data = dataset) %>%
        TukeyHSD(.)
    aov_test <- aov_test$condition %>%
        as.data.frame() %>%
        rownames_to_column("comparison") %>%
        as_tibble() %>%
        filter(grepl(cmp_group, comparison)) %>%
        mutate(comparison = str_remove(comparison, paste0("-*", cmp_group, "-*"))) %>%
        rename(p_adj = `p adj`) %>%
        filter(p_adj <= 0.05)

    obj_name <- paste0(cell_line, "_", data_type)
    assign(obj_name, dataset)
    saveRDS(obj_name, paste0("data/robj/", obj_name, ".rds"))

    plot_name <-
        ggplot(dataset, aes(condition, cells)) +
        geom_beeswarm(
            size = 2,
            color = "gray65",
            cex = 2
        ) +
        stat_summary(
            fun.data = mean_se,
            geom = "errorbar", 
            width = 0.3, 
            size = 1
        ) +
        stat_summary(
            fun = mean, 
            geom = "crossbar",
            width = 0.5,
            size = 0.5
        ) +
        labs(x = "",
             y = "Cells",
             title = cell_line) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    if (nrow(aov_test) > 0) {
        plot_name <- plot_name +
            geom_text(data = aov_test,
                      aes(x = comparison,
                          y = max(dataset$cells) * 1.1,
                          label = paste0("p=", sprintf("%.2e", p_adj))),
                      size = 3)
    }

    ggsave(paste0("data/output/figures/figure_4_",
                  cell_line,
                  "_",
                  data_type,
                  ".png"),
           width = 5,
           height = 5,
           plot = plot_name)
    print(plot_name)

    rm(list = obj_name, plot_name)
}
```







## Human tissue in lung - amputation studies data

```{r plotHsTissue, eval=TRUE}
label_table <- tribble(
    ~label, ~table_name,
    "Primary vs None", "Primary vs None",
    "Amputation vs None", "Amputation vs None"
)

file_name <- "../Source Data/Primary vs None.xml"
perc_human_cutoff <- 0.1

level_orders <- list(
    `Primary vs None` = c("Primary", "No Primary"),
    `Amputation vs None` = c("No Amp, Early Euth", "Early Amp, Late Euth", "No Amp, Late Euth")
)

for (i in seq_len(nrow(label_table))) {
    label <- label_table$label[i]
    table_name <- label_table$table_name[i]

    dataset <-
        as_tibble(read_pzfx(file_name, table = table_name)) %>%
        pivot_longer(
            cols = everything(),
            names_to = "condition",
            values_to = "perc_human"
        ) %>%
        filter(!is.na(perc_human)) %>%
        mutate(condition = factor(condition, levels = level_orders[[label]]))

    n_greater_table <-
        dataset %>%
        group_by(condition) %>%
        summarize(value = paste0(
            n_greater = sum(perc_human > perc_human_cutoff),
            "/",
            n()
            ))

    dataset_p <-
        pairwise.t.test(dataset$perc_human,
                        dataset$condition,
                        p.adjust.method = "bonf")$p.value %>%
        as.data.frame() %>%
        rownames_to_column("group_a") %>%
        rename(p = level_orders[[label]][1]) %>%
        select(group_a, p) %>%
        mutate(p = format(round(p, 3), nsmall = 3))

    ggplot(dataset,
           aes(x = condition, y = perc_human)) +
        geom_boxplot(outlier.alpha = 0) +
        geom_beeswarm(shape = 15,
                      cex = 3) +
        geom_hline(yintercept = perc_human_cutoff,
                   color = "gray",
                   linetype = "dashed") +
        geom_text(data = n_greater_table,
                  aes(x = condition,
                      y = max(dataset$perc_human * 1.1),
                      label = value)) +
        geom_text(data = dataset_p,
                  aes(x = group_a,
                      y = max(dataset$perc_human * 1.05),
                      label = paste0("p=", p)),
                  size = 3) +
        labs(x = "",
             y = "% Human Tissue\n(in lung)",
             title = label) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    ggsave(paste0("data/output/figures/figure_2_",
                  str_replace_all(label, " ", "_"),
                  ".png"),
           width = 4,
           height = 5)

}

```






# Conclusions

placeholder
 
