---
title: "Tumor-osteosarcoma interactions during metastatic colonization of the lung"
author: "Ryan Roberts, Matthew Cannon, James Reinecke"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        embed-resources: true
        self_contained: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        cache.vars: ""
        warning: false
        message: false
lightbox: true
---

```{r setup}
#| include: FALSE
library(tidyverse)
library(rrrSingleCellUtils)
library(Seurat)
library(stashPlot)

library(future)
library(future.apply)
if(Sys.info()['sysname'] == "Windows") {
  plan(multisession, workers = 5)
} else {
  plan(multicore)
}
options(future.globals.maxSize = 10000 * 1024^2)
```


## Check required software is available
```{r h5_check}
#| cache: false
hdf5_check()

seurat_version_check()
```


# Read in raw data and process to Seurat objects
The metadata for all samples is stored in misc/sample_metadata.tsv.

Note that in the sample_metadata, for samples with multiple references, the cutoff format is "GRCh38-40000;mm10---20000" where the species_pattern is at the beginning of the string, followed by the cutoff value (no dash or anything). This is split from the second cutoff value by a semicolon. If there is only one reference, just use the cutoff value, no species pattern is needed.

{{< include child_qmds/read_raw_data.qmd >}}

# Merge the data by species and make DimPlots

{{< include child_qmds/merge_dimplots.qmd >}}

# Run differential expression and GSEA

{{< include child_qmds/degs_gseaplots.qmd >}}


# Session info
```{r session_info}
sessionInfo()
```
