---
title: "Osteosarcoma - lung epithelial interactions"
author: "Ryan Roberts"
date: March 2022
output: knitrBootstrap::bootstrap_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.height = 10,
                      fig.width = 20,
                      dpi = 300,
                      bootstrap.show.code = FALSE,
                      bootstrap.show.output = FALSE,
                      bootstrap.show.message = FALSE)
```



# Osteosarcoma - lung epithelial interactions

## Goals:

Generate figures for the tumor-epithelial paracrine signaling, self-seeding mechanisms paper

## Samples:

placeholder

## Approach

placeholder

# Workflow

## Load libraries
```{r libraries, cache=FALSE, eval=TRUE}
library(tidyverse)
library(ggpubr)
library(rrrSingleCellUtils)
library(Seurat)
library(pzfx)
library(ggbeeswarm)
library(patchwork)
```

## Helper functions
```{r}
process_seurat <- function(obj, ...) {
    obj %>%
        NormalizeData(verbose = FALSE) %>%
        FindVariableFeatures(verbose = FALSE) %>%
        ScaleData(verbose = FALSE) %>%
        RunPCA(verbose = FALSE) %>%
        RunUMAP(dims = 1:20, verbose = FALSE) %>%
        FindNeighbors(verbose = FALSE) %>%
        FindClusters(verbose = FALSE, resolution = 0.4, ...)
}
theme_set(theme_pubr())

plot_cols <- c("#D43F3AFF", "#EEA236FF", "#357EBDFF", "#5CB85CFF", "#B8B8B8FF",
  "#9632B8FF",  "#46B8DAFF", "#90302DFF", "#A66D04FF", "#2D577FFF", "#3E7E3EFF",
  "#7D7D7DFF", "#6D1D87FF", "#097F9AFF", "#FF6E6AFF", "#FFBB70FF", "#68A4E3FF",
  "#79D379FF", "#CDCDCDFF", "#BF6BE2FF", "#69D1F3FF")

rDimPlot <- function(object, title = "Unlabeled Plot", subtitle = NULL,
  label = T, pt.size = 1, ...) {
  if(length(levels(Idents(object))) < 22) {
    p <- Seurat::DimPlot(
      object = object, 
      label = label, 
      pt.size = pt.size,
      cols = alpha(plot_cols, 0.6),
      ...) +
      labs(title = title, subtitle = subtitle) +
      theme(legend.position = "none") +
      coord_fixed()
    return(p)
  } else {
    print("Too many identity classes to use this function. Requires <22.")
  }
}

rFeaturePlot <- function(object, features, title = "Unlabeled Plot", subtitle = NULL,
  pt.size = 1, order = T, ...) {
  p <- Seurat::FeaturePlot(
    object = object,
    features = features,
    pt.size = pt.size,
    order = order,
    cols = (c("gray95", "#1F00B8")),
    ...) +
    labs(title = title, subtitle = subtitle) +
    theme(plot.title = element_text(hjust = 0)) +
    coord_fixed()
  return(p)
}

set.seed(888)
```

## Load and organize co-culture data
```{r , eval=TRUE}
scrna_path <- "D:/OneDrive - Nationwide Children's Hospital/BIScratch/"
deets_cocx <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~src,
    "S0001", 16000, 14, "mono-cx",
    "S0005", 16000, 14, "co-cx")

cocx <- lapply(seq_len(nrow(deets_cocx)), function(i) {
    s <- tenx_load_qc(paste0(scrna_path,
                deets_cocx$id[i],
                "/filtered_feature_bc_matrix/"),
            violin_plot = FALSE) %>%
        subset(
            nCount_RNA < deets_cocx$max_count[i] &
            percent.mt < deets_cocx$max_percent_mt[i]) %>%
        process_seurat()
})
names(cocx) <- deets_cocx$id
for(i in seq_along(cocx)) {
    cocx[[i]]$src <- deets_cocx$src[i]
}
```

## Intercellular signaling analysis on co-culture data
Which signals produced by lung epithelial cells elicit the transcriptional changes that we see in osteosarcoma cells in co-culture?
In NicheNET, define the target gene set by DEG analysis comparing osteosarcoma cells in co-culture to cultured alone.
Define senders as the epithelial cells and receivers as the osteosarcoma cells.

```{r, eval=TRUE}
fig_nichenet <- list()

cx <- merge(
    cocx$S0001,
    y = cocx$S0005,
    add.cell.ids = c("monocx", "cocx"),
    project = "InterSignaling")

cx <- process_seurat(cx)

fig_nichenet$umap_src <- rDimPlot(cx, 
    group.by = "src",
    title = "OS-17 + HBEC, source")
fig_nichenet$umap <- rDimPlot(cx, title = "OS-17 + HBEC, clusters")
fig_nichenet$krt19 <- rFeaturePlot(cx, features = "KRT19", title = "KRT19")
fig_nichenet$col1a1 <- rFeaturePlot(cx, features = "COL1A1", title = "COL1A1")

# Suspect that cluster 9 is a group of doublets, remove
cx <- subset(cx, idents = c(8, 9), invert = T)

cx <- RenameIdents(cx,
    `4` = "OS-MonoCx",
    `1` = "OS-CoCx", `7` = "OS-CoCx",
    `0` = "HBEC-CoCx", `2` = "HBEC-CoCx", `5` = "HBEC-CoCx",
    `3` = "HBEC-MonoCx", `6` = "HBEC-MonoCx")

fig_nichenet$assign <- rDimPlot(cx, title = "OS-17 + HBEC, assignment")

target_genes <- find_tar_genes(cx,
    id1 = "OS-CoCx",
    id2 = "OS-MonoCx",
    logfc = 0.5)

LRT_analysis_1 <- find_ligands(cx,
    target_genes,
    senders = c("HBEC-CoCx", "HBEC-MonoCx"),
    receiver = "OS-CoCx",
    rec_pct = 0.07,
    stringency = "strict")

fig_nichenet$LT_plot <- plot_complex_heatmap(LRT_analysis_1,
    grid_color_high = "#004466")
fig_nichenet$LR_plot <- LRT_analysis_1[[3]] +
    scale_fill_gradient(low = "gray98", high = "#006600")
```

## Assemble the panel for NicheNET analysis
```{r}
p1 <- fig_nichenet$umap + fig_nichenet$umap_src + fig_nichenet$krt19 + 
    fig_nichenet$col1a1 + plot_layout(ncol = 4)
p2 <- fig_nichenet$assign + fig_nichenet$LT_plot + plot_layout(widths = c(1, 3))
nn_panel <- p1 / p2 / fig_nichenet$LR_plot
nn_panel
```

## Isolate the co-culture tumor and evaluate IL6 and CXCL8 expression
```{r}
tumor_cocx <- subset(cx, idents = "OS-CoCx")

# Remove the doublets that express epithelial markers
tumor_cocx <- subset(tumor_cocx, cells = readRDS("doublets.rds"), invert = T)

tumor_cocx <- process_seurat(tumor_cocx)
tumor_cocx <- RunUMAP(tumor_cocx,
        dims = 1:20,
        n.neighbors = 20,
        min.dist = 0.1,
        local.connectivity = 5,
        spread = 1,
        verbose = FALSE) %>%
    FindNeighbors(verbose = FALSE) %>%
    FindClusters(verbose = FALSE, resolution = 0.25)

p1 <- rDimPlot(tumor_cocx,
    title = "Tumor in Co-culture")
p2 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = IL6 > 3),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6 Expression >3") +
    theme(legend.position = "none") +
    coord_fixed()
p3 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = CXCL8 > 6),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CXCL8 Expression >6") +
    theme(legend.position = "none") +
    coord_fixed()
p4 <- rFeaturePlot(tumor_cocx,
    "nCount_RNA",
    title = "Total Counts",
    max.cutoff = 12000)
tumor_sub_panel <- p1 | p2 | p3 | p4
tumor_sub_panel
```

## Load human met scRNA-seq and create Seurat objects

```{r}
deets_mets <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~src,
    "S0058", 50000, 14, "met-1",
    "S0059", 50000, 14, "met-2")

mets <- lapply(seq_len(nrow(deets_mets)), function(i) {
    s <- tenx_load_qc(paste0(scrna_path,
                deets_mets$id[i],
                "/filtered_feature_bc_matrix/"),
            violin_plot = FALSE) %>%
        subset(
            nCount_RNA < deets_mets$max_count[i] &
            percent.mt < deets_mets$max_percent_mt[i]) %>%
        process_seurat()
})

names(mets) <- deets_mets$id
for(i in seq_along(mets)) {
    mets[[i]]$src <- deets_mets$src[i]
}
```

## Check for co-occurrence of IL6, CXCL8, p21, and Ki67 in human lung mets

```{r}
rDimPlot(mets$S0058, title = "Clusters")
rFeaturePlot(mets$S0058, "COL1A1", title = "COL1A1 Expression")
DimPlot(mets$S0058,
        cells.highlight = WhichCells(mets$S0058, expression = IL6 > 2),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6 Expression >2") +
    theme(legend.position = "none") +
    coord_fixed()

# Isolate the tumor cells (list previously generated)
mets$tS0058 <- subset(mets$S0058, cells = readRDS("S0058TumorCells.rds"))

mets$tS0058 <- process_seurat(mets$tS0058)
rDimPlot(mets$tS0058, title = "Tumor Subclusters")
rFeaturePlot(mets$tS0058, "IL6", title = "IL6 Expression")
rFeaturePlot(mets$tS0058, "CXCL8", title = "CXCL8 Expression")
DimPlot(mets$tS0058,
        cells.highlight = WhichCells(mets$tS0058, expression = IL6 > 1),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6 Expression >1") +
    theme(legend.position = "none") +
    coord_fixed()




```


















## Validation of IL1 as the primary lung-to-tumor signaling

```{r}

```






## Figure 2: Human tissue in lung
```{r plotHsTissue, eval=TRUE}
label_table <- tribble(
    ~label, ~table_name,
    "Primary vs None", "Primary vs None",
    "Amputation vs None", "Amputation vs None"
)

file_name <- "../Source Data/Primary vs None.xml"
perc_human_cutoff <- 0.1

level_orders <- list(
    `Primary vs None` = c("Primary", "No Primary"),
    `Amputation vs None` = c("No Amp, Early Euth", "Early Amp, Late Euth", "No Amp, Late Euth")
)

for (i in seq_len(nrow(label_table))) {
    label <- label_table$label[i]
    table_name <- label_table$table_name[i]

    dataset <-
        as_tibble(read_pzfx(file_name, table = table_name)) %>%
        pivot_longer(
            cols = everything(),
            names_to = "condition",
            values_to = "perc_human"
        ) %>%
        filter(!is.na(perc_human)) %>%
        mutate(condition = factor(condition, levels = level_orders[[label]]))

    n_greater_table <-
        dataset %>%
        group_by(condition) %>%
        summarize(value = paste0(
            n_greater = sum(perc_human > perc_human_cutoff),
            "/",
            n()
            ))

    dataset_p <-
        pairwise.t.test(dataset$perc_human,
                        dataset$condition,
                        p.adjust.method = "bonf")$p.value %>%
        as.data.frame() %>%
        rownames_to_column("group_a") %>%
        rename(p = level_orders[[label]][1]) %>%
        select(group_a, p) %>%
        mutate(p = format(round(p, 3), nsmall = 3))

    ggplot(dataset,
           aes(x = condition, y = perc_human)) +
        geom_boxplot(outlier.alpha = 0) +
        geom_beeswarm(shape = 15,
                      cex = 3) +
        geom_hline(yintercept = perc_human_cutoff,
                   color = "gray",
                   linetype = "dashed") +
        geom_text(data = n_greater_table,
                  aes(x = condition,
                      y = max(dataset$perc_human * 1.1),
                      label = value)) +
        geom_text(data = dataset_p,
                  aes(x = group_a,
                      y = max(dataset$perc_human * 1.05),
                      label = paste0("p=", p)),
                  size = 3) +
        labs(x = "",
             y = "% Human Tissue\n(in lung)",
             title = label) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    ggsave(paste0("data/output/figures/figure_2_",
                  str_replace_all(label, " ", "_"),
                  ".png"),
           width = 4,
           height = 5)

}

```


## Get list of data sources
```{r}
data_sources <- readxl::read_excel("D:/OneDrive - Nationwide Children's Hospital/Papers/Active Manuscripts/Self seeding and IL6-IL1 paracrine loop/Source Data/Locations and Names of Data.xlsx")
```

## Figure 4: Import transwell data and create data object
Save data objects with metadata to data/
Compare to media only

```{r plotCytokines, eval=TRUE}
paper_path <- "../Source Data/"

migration <- list()

migration$cytokine <- list()
migration$cond_media <- list()

migration$cytokine$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 Migration", "metastatic", "to_cytokine",
    "143B", "143B Migration", "metastatic", "to_cytokine",
    "K7M2", "K7M2 Migration", "metastatic", "to_cytokine",
    "OS-25", "OS-25 Migration", "non-metastatic", "to_cytokine",
    "OSCA8", "OSCA8 Migration", "", "to_cytokine")

migration$cond_media$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 migration assay (6.14.18) to inhibitors",
        "metastatic", "to_inhibitors",
    "143B", "143B migration (2018-06-01) to inhibitors",
        "metastatic", "to_inhibitors",
    "K7M2", "K7M2 migration (2018-06-26) to inhibitors",
        "metastatic", "to_inhibitors",
    "SaOS2", "SaOS2 migration (2018-07-11) to inhibitors",
        "non-metastatic", "to_inhibitors",
    "MG63", "MG-63 migration (#1 2018-05-25) to inhibitors",
        "non-metastatic", "to_inhibitors")

migration$cytokine$data <- lapply(
    migration$cytokine$source$table_name, 
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(
            description = x,
            met_potential = migration$cytokine$source$met_potential[
                migration$cytokine$source$table_name == x]) %>%
        filter(condition != "CXCL12" & condition != "OSM") %>%
        mutate(condition =
            str_replace(condition, "^Media$", "Media") %>%
            str_replace("1% FBS", "FBS")) %>%
        mutate(condition = factor(condition, 
            levels = c("Media", "FBS", "IL6", "IL8")))
    })
names(migration$cytokine$data) <- migration$cytokine$source$cell_line

migration$cond_media$data <- lapply(
    migration$cond_media$source$table_name,
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(description = x,
            met_potential = migration$cond_media$source$met_potential[
                migration$cond_media$source$table_name == x])
        ))
)
        if (data_type == "to_cytokine") {
            dataset <- dataset %>%
            
    } else if (data_type == "to_inhibitors") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("Cond Media", "Cond. media (C.M.)") %>%
                       str_replace("CM \\+ sc144", "C.M. + SC144") %>%
                       str_replace("CM \\+ ladarixin", "C.M. + DF2156A") %>%
                       str_replace("CM \\+ both", "C.M. + both")) %>%
            mutate(condition = factor(condition, levels = c('Media only',
                                                            '1% FBS',
                                                            'Cond. media (C.M.)',
                                                            'C.M. + SC144',
                                                            'C.M. + DF2156A',
                                                            'C.M. + both')))
    } else if (data_type == "to_conditioned_media") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("1% FBS", "Media + FBS") %>%
                       str_replace("Cond Media", "Cond. media")) %>%
            mutate(condition = factor(condition, levels = c("Media only",
                                                            "Media + FBS",
                                                            "Cond. media")))
    }

    cmp_group <- if_else(data_type == "to_conditioned_media",
                         true = "Media . FBS",
                         false = "Media only")
    aov_test <- aov(cells ~ condition, data = dataset) %>%
        TukeyHSD(.)
    aov_test <- aov_test$condition %>%
        as.data.frame() %>%
        rownames_to_column("comparison") %>%
        as_tibble() %>%
        filter(grepl(cmp_group, comparison)) %>%
        mutate(comparison = str_remove(comparison, paste0("-*", cmp_group, "-*"))) %>%
        rename(p_adj = `p adj`) %>%
        filter(p_adj <= 0.05)

    obj_name <- paste0(cell_line, "_", data_type)
    assign(obj_name, dataset)
    saveRDS(obj_name, paste0("data/robj/", obj_name, ".rds"))

    plot_name <-
        ggplot(dataset, aes(condition, cells)) +
        geom_beeswarm(
            size = 2,
            color = "gray65",
            cex = 2
        ) +
        stat_summary(
            fun.data = mean_se,
            geom = "errorbar", 
            width = 0.3, 
            size = 1
        ) +
        stat_summary(
            fun = mean, 
            geom = "crossbar",
            width = 0.5,
            size = 0.5
        ) +
        labs(x = "",
             y = "Cells",
             title = cell_line) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    if (nrow(aov_test) > 0) {
        plot_name <- plot_name +
            geom_text(data = aov_test,
                      aes(x = comparison,
                          y = max(dataset$cells) * 1.1,
                          label = paste0("p=", sprintf("%.2e", p_adj))),
                      size = 3)
    }

    ggsave(paste0("data/output/figures/figure_4_",
                  cell_line,
                  "_",
                  data_type,
                  ".png"),
           width = 5,
           height = 5,
           plot = plot_name)
    print(plot_name)

    rm(list = obj_name, plot_name)
}
```





# Conclusions

placeholder
 
