---
title: "Osteosarcoma - lung epithelial interactions"
author: "Ryan Roberts"
date: March 2022
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.height = 10,
                      fig.width = 20,
                      dpi = 300,
                      bootstrap.show.code = FALSE,
                      bootstrap.show.output = FALSE,
                      bootstrap.show.message = FALSE)
```



# Osteosarcoma - lung epithelial interactions

## Goals:

Generate figures for the tumor-epithelial paracrine signaling, self-seeding mechanisms paper

## Samples:

sample	type	primary loc
BC2	primary	femur
BC3	primary	tibia
BC5	primary	femur
BC6	primary	ulna
BC10	lung met	femur
BC11	recurrent	femur
BC16	primary	tibia
BC17	lung met	tibia
BC20	recurrent	femur
BC21	primary	femur
BC22	primary	femur


# Workflow

## Load libraries
```{r libraries, cache=FALSE, eval=TRUE}
library(tidyverse)
library(ggpubr)
library(rrrSingleCellUtils)
library(Seurat)
library(pzfx)
library(ggbeeswarm)
library(patchwork)
```

## Helper functions
```{r}
set.seed(888)
```

## Load and organize co-culture data
```{r , eval=TRUE}
scrna_path <- "/home/gdrobertslab/lab/Counts/"
deets_cocx <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~file,
    "S0001", 16000, 14, "S0001-mix/outs",
    "S0005", 16000, 14, "S0005-CoCx-2/outs")

cocx <- lapply(seq_len(nrow(deets_cocx)), function(i) {
    test <- tenx_load_qc(paste0(scrna_path,
                        deets_cocx$file[i],
                        "/filtered_feature_bc_matrix/"),
                 violin_plot = FALSE) %>%
        subset(nCount_RNA < deets_cocx$max_count[i] &
                   percent.mt < deets_cocx$max_percent_mt[i]) %>%
        process_seurat()
})
names(cocx) <- deets_cocx$id
for (i in seq_along(cocx)) {
    cocx[[i]]$src <- deets_cocx$id[i]
}
```

## Intercellular signaling analysis on co-culture data
Which signals produced by lung epithelial cells elicit the transcriptional changes that we see in osteosarcoma cells in co-culture?
In NicheNET, define the target gene set by DEG analysis comparing osteosarcoma cells in co-culture to cultured alone.
Define senders as the epithelial cells and receivers as the osteosarcoma cells.

```{r, eval=TRUE}
fig_nichenet <- list()

cx <- merge(
    cocx$S0001,
    y = cocx$S0005,
    add.cell.ids = c("monocx", "cocx"),
    project = "InterSignaling") %>%
    JoinLayers()

opt_res <-
    optimize_silhouette(cx, summary_plot = FALSE) %>%
    arrange(desc(sil_vals)) %>%
    slice_head(n = 1) %>%
    pull(res_vals)

cx <- process_seurat(cx, resolution  = opt_res)

fig_nichenet$umap_src <- r_dim_plot(cx,
    group.by = "src",
    title = "OS-17 + HBEC, source")
fig_nichenet$umap <- r_dim_plot(cx, title = "OS-17 + HBEC, clusters")
fig_nichenet$krt19 <- r_feature_plot(cx, features = "KRT19", title = "KRT19") # HBEC marker
fig_nichenet$col1a1 <- r_feature_plot(cx, features = "COL1A1", title = "COL1A1") # OS marker

offset_value <- 0.0001
marker_data <-
    GetAssayData(
        cx
    ) %>%
    t() %>%
    as.data.frame() %>%
    select(KRT19, COL1A1) %>%
    mutate(
        KRT19 = KRT19 + offset_value,
        COL1A1 = COL1A1 + offset_value,
        marker_ratio = KRT19 / COL1A1
    )

cx <-
    AddMetaData(
        cx,
        marker_data
    )

r_feature_plot(cx, "marker_ratio", title = "KRT19/COL1A1 Ratio")
feature_hist(cx, features = "marker_ratio")

ratio_threshold <- 1000
tumor_clusters <-
    cx@meta.data %>%
    group_by(seurat_clusters, src) %>%
    summarize(
        mean_ratio = mean(marker_ratio),
        .groups = "drop"
    ) %>%
    mutate(
        is_tumor = if_else(
            mean_ratio < ratio_threshold,
            "OS",
            "HBEC"
        ),
        culture = if_else(
            src == "S0001",
            "MonoCx",
            "CoCx"
        ),
        group = paste0(is_tumor, "-", culture)
    )

cx <-
    AddMetaData(
        cx,
        cx@meta.data %>%
        rownames_to_column("cell") %>%
        select("cell", "seurat_clusters", "src") %>%
        merge(tumor_clusters, by = c("seurat_clusters", "src")) %>%
        column_to_rownames("cell")
    )

r_dim_plot(cx, group.by = c("seurat_clusters", "group", "src")) /
    r_feature_plot(cx, c("KRT19", "COL1A1"))

Idents(cx) <- cx$group

qs::qsave(cx, "output/rdata/cx.qs")

# Suspect that cluster 9 is a group of doublets, remove
# cx <- subset(cx, idents = c(8, 9), invert = T)

# cx <- RenameIdents(cx,
#     `4` = "OS-MonoCx",
#     `1` = "OS-CoCx", `7` = "OS-CoCx",
#     `0` = "HBEC-CoCx", `2` = "HBEC-CoCx", `5` = "HBEC-CoCx",
#     `3` = "HBEC-MonoCx", `6` = "HBEC-MonoCx")

fig_nichenet$assign <- r_dim_plot(cx, title = "OS-17 + HBEC, assignment")

target_genes <- find_tar_genes(cx,
    id1 = "OS-CoCx",
    id2 = "OS-MonoCx",
    logfc = 0.5)

LRT_analysis_1 <- find_ligands(cx,
    target_genes,
    senders = c("HBEC-CoCx", "HBEC-MonoCx"),
    receiver = "OS-CoCx",
    rec_pct = 0.07,
    stringency = "strict")

fig_nichenet$LT_plot <- plot_complex_heatmap(LRT_analysis_1,
    grid_color_high = "#004466")
fig_nichenet$LR_plot <- LRT_analysis_1[[3]] +
    scale_fill_gradient(low = "gray98", high = "#006600")
```

## Assemble the panel for NicheNET analysis
```{r}
p1 <- fig_nichenet$umap + fig_nichenet$umap_src + fig_nichenet$krt19 +
    fig_nichenet$col1a1 + plot_layout(ncol = 4)
p2 <- fig_nichenet$assign + fig_nichenet$LT_plot + plot_layout(widths = c(1, 3))
nn_panel <- p1 / p2 / fig_nichenet$LR_plot
# Figure 3B/C/D/E
nn_panel
```

## Isolate the co-culture tumor and evaluate IL6 and CXCL8 expression
```{r}
tumor_cocx <- subset(cx, idents = "OS-CoCx")

# Remove the doublets that express epithelial markers
tumor_cocx <- subset(tumor_cocx, cells = readRDS("doublets.rds"), invert = T)

tumor_cocx <- process_seurat(tumor_cocx)
tumor_cocx <- RunUMAP(tumor_cocx,
        dims = 1:20,
        n.neighbors = 20,
        min.dist = 0.1,
        local.connectivity = 5,
        spread = 1,
        verbose = FALSE) %>%
    FindNeighbors(verbose = FALSE) %>%
    FindClusters(verbose = FALSE, resolution = 0.25)

p1 <- r_dim_plot(tumor_cocx,
    title = "Tumor in Co-culture")
p2 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = IL6 > 3),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6 Expression >3") +
    theme(legend.position = "none") +
    coord_fixed()
p3 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = CXCL8 > 6),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CXCL8 Expression >6") +
    theme(legend.position = "none") +
    coord_fixed()
p4 <- r_feature_plot(tumor_cocx,
    "nCount_RNA",
    title = "Total Counts",
    max.cutoff = 12000)
tumor_sub_panel <- p1 | p2 | p3 | p4
# Figure 4A
tumor_sub_panel
```

## Load human met scRNA-seq and create Seurat objects

```{r}
deets_mets <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~src,
    "S0058", 50000, 14, "met-1",
    "S0059", 50000, 14, "met-2")

mets <- lapply(seq_len(nrow(deets_mets)), function(i) {
    s <- tenx_load_qc(paste0(scrna_path,
                deets_mets$id[i],
                "/filtered_feature_bc_matrix/"),
            violin_plot = FALSE) %>%
        subset(
            nCount_RNA < deets_mets$max_count[i] &
            percent.mt < deets_mets$max_percent_mt[i]) %>%
        process_seurat()
})

names(mets) <- deets_mets$id
for (i in seq_along(mets)) {
    mets[[i]]$src <- deets_mets$src[i]
}

tumor_cells <- readRDS("tumor_cells.rds")

mets <- merge(mets[[1]], mets[[2:length(mets)]],
    add.cell.ids = deets_mets$id) %>%
    process_seurat()
mets$tumor <- "stroma"
mets$tumor[names(mets$tumor) %in% tumor_cells] <- "tumor"

r_dim_plot(mets, "Merged human lung metastases")
r_dim_plot(mets, group.by = "tumor", "Cells identified as tumor")

mets_tumor <- subset(mets, tumor == "tumor") %>%
    process_seurat()

p1 <- r_dim_plot(mets_tumor, "Isolated tumor cells")
p2 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = IL6 > 0),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6") +
    theme(legend.position = "none") +
    coord_fixed()
p3 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = CXCL8 > 1),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CXCL8") +
    theme(legend.position = "none") +
    coord_fixed()
p4 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = MKI67 > 1),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "MKI67") +
    theme(legend.position = "none") +
    coord_fixed()
p5 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = CDKN1A > 1.5),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CDKN1A") +
    theme(legend.position = "none") +
    coord_fixed()
p1 | p2 | p3 | p4 | p5
```

## Check for co-occurrence of IL6, CXCL8, p21, and Ki67 in human lung mets

```{r}
tmets <- list()
for (i in c("S0058", "S0059")) {
    r_dim_plot(mets[[i]], title = paste(i, "Clusters"))
    r_feature_plot(mets[[i]], "IL6", title = paste(i, "COL1A1 Expression"))
    DimPlot(mets[[i]],
            cells.highlight = WhichCells(mets[[i]], expression = IL6 > 1),
            cols.highlight = "#1F00B8",
            pt.size = 1) +
        labs(title = paste(i, "IL6 Expression >2")) +
        theme(legend.position = "none") +
        coord_fixed()

    # Isolate the tumor cells (list previously generated)
    # tmets[[i]] <- subset(mets[[i]],
    #     cells = readRDS(paste0(i, "TumorCells.rds")))

    # tmets[[i]] <- process_seurat(tmets[[i]])
    tmets[[i]] <- mets[[i]]
    r_dim_plot(tmets[[i]], title = paste(i, "Tumor Subclusters"))
    r_feature_plot(tmets[[i]], "IL6", title = paste(i, "IL6 Expression"))
    r_feature_plot(tmets[[i]], "CXCL8", title = paste(i, "CXCL8 Expression"))
    DimPlot(tmets[[i]],
            cells.highlight = WhichCells(tmets[[i]], expression = IL6 > 1),
            cols.highlight = "#1F00B8",
            pt.size = 1) +
        labs(title = paste(i, "IL6 Expression >1")) +
        theme(legend.position = "none") +
        coord_fixed()
}

tumor_cocx <- SCTransform(tumor_cocx)
s <- c("counts", "data", "scale.data")
p <- lapply(s, function(i) {
    t <- GetAssayData(tumor_cocx, slot = i) %>%
        as.data.frame() %>%
        t() %>%
        as.data.frame()

    return(
        ggplot(t, aes(IL6, CDKN1A)) +
            geom_point()
    )
})
p[[1]] | p[[2]] | p[[3]]

```


















## Validation of IL1 as the primary lung-to-tumor signaling

```{r}

```






## Figure 2: Human tissue in lung
```{r plotHsTissue, eval=TRUE}
label_table <- tribble(
    ~label, ~table_name,
    "Primary vs None", "Primary vs None",
    "Amputation vs None", "Amputation vs None"
)

file_name <- "../Source Data/Primary vs None.xml"
perc_human_cutoff <- 0.1

level_orders <- list(
    `Primary vs None` = c("Primary", "No Primary"),
    `Amputation vs None` = c("No Amp, Early Euth", "Early Amp, Late Euth", "No Amp, Late Euth")
)

for (i in seq_len(nrow(label_table))) {
    label <- label_table$label[i]
    table_name <- label_table$table_name[i]

    dataset <-
        as_tibble(read_pzfx(file_name, table = table_name)) %>%
        pivot_longer(
            cols = everything(),
            names_to = "condition",
            values_to = "perc_human"
        ) %>%
        filter(!is.na(perc_human)) %>%
        mutate(condition = factor(condition, levels = level_orders[[label]]))

    n_greater_table <-
        dataset %>%
        group_by(condition) %>%
        summarize(value = paste0(
            n_greater = sum(perc_human > perc_human_cutoff),
            "/",
            n()
            ))

    dataset_p <-
        pairwise.t.test(dataset$perc_human,
                        dataset$condition,
                        p.adjust.method = "bonf")$p.value %>%
        as.data.frame() %>%
        rownames_to_column("group_a") %>%
        rename(p = level_orders[[label]][1]) %>%
        select(group_a, p) %>%
        mutate(p = format(round(p, 3), nsmall = 3))

    ggplot(dataset,
           aes(x = condition, y = perc_human)) +
        geom_boxplot(outlier.alpha = 0) +
        geom_beeswarm(shape = 15,
                      cex = 3) +
        geom_hline(yintercept = perc_human_cutoff,
                   color = "gray",
                   linetype = "dashed") +
        geom_text(data = n_greater_table,
                  aes(x = condition,
                      y = max(dataset$perc_human * 1.1),
                      label = value)) +
        geom_text(data = dataset_p,
                  aes(x = group_a,
                      y = max(dataset$perc_human * 1.05),
                      label = paste0("p=", p)),
                  size = 3) +
        labs(x = "",
             y = "% Human Tissue\n(in lung)",
             title = label) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    ggsave(paste0("data/output/figures/figure_2_",
                  str_replace_all(label, " ", "_"),
                  ".png"),
           width = 4,
           height = 5)

}

```


## Get list of data sources
```{r}
data_sources <- readxl::read_excel("D:/OneDrive - Nationwide Children's Hospital/Papers/Active Manuscripts/Self seeding and IL6-IL1 paracrine loop/Source Data/Locations and Names of Data.xlsx")
```

## Figure 4: Import transwell data and create data object
Save data objects with metadata to data/
Compare to media only

```{r plotCytokines, eval=TRUE}
paper_path <- "../Source Data/"

migration <- list()

migration$cytokine <- list()
migration$cond_media <- list()

migration$cytokine$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 Migration", "metastatic", "to_cytokine",
    "143B", "143B Migration", "metastatic", "to_cytokine",
    "K7M2", "K7M2 Migration", "metastatic", "to_cytokine",
    "OS-25", "OS-25 Migration", "non-metastatic", "to_cytokine",
    "OSCA8", "OSCA8 Migration", "", "to_cytokine")

migration$cond_media$source <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 migration assay (6.14.18) to inhibitors",
        "metastatic", "to_inhibitors",
    "143B", "143B migration (2018-06-01) to inhibitors",
        "metastatic", "to_inhibitors",
    "K7M2", "K7M2 migration (2018-06-26) to inhibitors",
        "metastatic", "to_inhibitors",
    "SaOS2", "SaOS2 migration (2018-07-11) to inhibitors",
        "non-metastatic", "to_inhibitors",
    "MG63", "MG-63 migration (#1 2018-05-25) to inhibitors",
        "non-metastatic", "to_inhibitors")

migration$cytokine$data <- lapply(
    migration$cytokine$source$table_name,
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(
            description = x,
            met_potential = migration$cytokine$source$met_potential[
                migration$cytokine$source$table_name == x]) %>%
        filter(condition != "CXCL12" & condition != "OSM") %>%
        mutate(condition =
            str_replace(condition, "^Media$", "Media") %>%
            str_replace("1% FBS", "FBS")) %>%
        mutate(condition = factor(condition,
            levels = c("Media", "FBS", "IL6", "IL8")))
    })
names(migration$cytokine$data) <- migration$cytokine$source$cell_line

migration$cond_media$data <- lapply(
    migration$cond_media$source$table_name,
    function(x) {
        as_tibble(read_pzfx(
            paste0(paper_path, "Figures for paper.xml"),
            table = x)) %>%
        gather("condition", "cells") %>%
        mutate(description = x,
            met_potential = migration$cond_media$source$met_potential[
                migration$cond_media$source$table_name == x])
        ))
)
        if (data_type == "to_cytokine") {
            dataset <- dataset %>%

    } else if (data_type == "to_inhibitors") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("Cond Media", "Cond. media (C.M.)") %>%
                       str_replace("CM \\+ sc144", "C.M. + SC144") %>%
                       str_replace("CM \\+ ladarixin", "C.M. + DF2156A") %>%
                       str_replace("CM \\+ both", "C.M. + both")) %>%
            mutate(condition = factor(condition, levels = c('Media only',
                                                            '1% FBS',
                                                            'Cond. media (C.M.)',
                                                            'C.M. + SC144',
                                                            'C.M. + DF2156A',
                                                            'C.M. + both')))
    } else if (data_type == "to_conditioned_media") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("1% FBS", "Media + FBS") %>%
                       str_replace("Cond Media", "Cond. media")) %>%
            mutate(condition = factor(condition, levels = c("Media only",
                                                            "Media + FBS",
                                                            "Cond. media")))
    }

    cmp_group <- if_else(data_type == "to_conditioned_media",
                         true = "Media . FBS",
                         false = "Media only")
    aov_test <- aov(cells ~ condition, data = dataset) %>%
        TukeyHSD(.)
    aov_test <- aov_test$condition %>%
        as.data.frame() %>%
        rownames_to_column("comparison") %>%
        as_tibble() %>%
        filter(grepl(cmp_group, comparison)) %>%
        mutate(comparison = str_remove(comparison, paste0("-*", cmp_group, "-*"))) %>%
        rename(p_adj = `p adj`) %>%
        filter(p_adj <= 0.05)

    obj_name <- paste0(cell_line, "_", data_type)
    assign(obj_name, dataset)
    saveRDS(obj_name, paste0("data/robj/", obj_name, ".rds"))

    plot_name <-
        ggplot(dataset, aes(condition, cells)) +
        geom_beeswarm(
            size = 2,
            color = "gray65",
            cex = 2
        ) +
        stat_summary(
            fun.data = mean_se,
            geom = "errorbar",
            width = 0.3,
            size = 1
        ) +
        stat_summary(
            fun = mean,
            geom = "crossbar",
            width = 0.5,
            size = 0.5
        ) +
        labs(x = "",
             y = "Cells",
             title = cell_line) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    if (nrow(aov_test) > 0) {
        plot_name <- plot_name +
            geom_text(data = aov_test,
                      aes(x = comparison,
                          y = max(dataset$cells) * 1.1,
                          label = paste0("p=", sprintf("%.2e", p_adj))),
                      size = 3)
    }

    ggsave(paste0("data/output/figures/figure_4_",
                  cell_line,
                  "_",
                  data_type,
                  ".png"),
           width = 5,
           height = 5,
           plot = plot_name)
    print(plot_name)

    rm(list = obj_name, plot_name)
}
```





# Conclusions

placeholder

