---
title: "Osteosarcoma - lung epithelial interactions"
author: "Ryan Roberts"
date: March 2022
output: knitrBootstrap::bootstrap_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.height = 10,
                      fig.width = 20,
                      dpi = 300,
                      bootstrap.show.code = FALSE,
                      bootstrap.show.output = FALSE,
                      bootstrap.show.message = FALSE)
```



# Osteosarcoma - lung epithelial interactions

## Goals:

Generate figures for the tumor-epithelial paracrine signaling, self-seeding mechanisms paper

## Samples:

placeholder

## Approach

placeholder

# Workflow

## Load libraries
```{r libraries, cache=FALSE, eval=TRUE}
library(tidyverse)
theme_set(theme_classic())
theme_update(plot.title = element_text(hjust = 0.5))
library(rrrSingleCellUtils)
library(Seurat)
library(pzfx)
library(ggbeeswarm)
```

## Load and organize co-culture data
```{r , eval=TRUE}
scrna_path <- "data/"
deets_cocx <- tribble(
    ~id, ~max_count, ~max_percent_mt,
    "S0001", 16000, 14,
    "S0005", 16000, 14)

cocx <- lapply(deets_cocx$id, function(i) {
    tenx_load_qc(paste(scrna_path,
                       i,
                       "/filtered_feature_bc_matrix/",
                       sep = ""),
                 violin_plot = FALSE)
})
names(cocx) <- deets_cocx$id
```

## Figure 4: Import and graph transwell data
Save data objects with metadata to data/
Compare to media only

cytokine 
condition media migration
```{r plotCytokines, eval=TRUE}
paper_path <- "../"
migration_tables <- tribble(
    ~cell_line, ~table_name, ~met_potential, ~type,
    "OS-17", "OS-17 Migration", "metastatic", "to_cytokine",
    "143B", "143B Migration", "metastatic", "to_cytokine",
    "K7M2", "K7M2 Migration", "metastatic", "to_cytokine",
    "OS-25", "OS-25 Migration", "non-metastatic", "to_cytokine",
    "OSCA8", "OSCA8 Migration", "", "to_cytokine",
    "OS-17", "OS-17 migration assay (6.14.18) to inhibitors", "metastatic", "to_inhibitors",
    "143B", "143B migration (2018-06-01) to inhibitors", "metastatic", "to_inhibitors",
    "K7M2", "K7M2 migration (2018-06-26) to inhibitors", "metastatic", "to_inhibitors",
    "SaOS2", "SaOS2 migration (2018-07-11) to inhibitors", "non-metastatic", "to_inhibitors",
    "MG63", "MG-63 migration (#1 2018-05-25) to inhibitors", "non-metastatic", "to_inhibitors",
    "OS-17", "OS-17 migration assay (6.14.18) to conditioned media", "metastatic", "to_conditioned_media",
    "143B", "143B migration (2018-06-01) to conditioned media", "metastatic", "to_conditioned_media",
    "K7M2", "K7M2 migration (2018-06-26) to conditioned media", "metastatic", "to_conditioned_media",
    "MG63", "MG-63 migration (#1 2018-05-25) to conditioned media", "non-metastatic", "to_conditioned_media",
    "SaOS2", "SaOS2 migration (2018-07-11) to conditioned media", "non-metastatic", "to_conditioned_media",
)
# get rid of SDF-1 (CXCL12) and Oncostatin M from "to_cytokine" datasets
# CXCL8 is IL8

for (i in seq_len(nrow(migration_tables))) {
    cell_line <- migration_tables$cell_line[i]
    table_name <- migration_tables$table_name[i]
    data_type <- migration_tables$type[i]

    dataset <-
        as_tibble(read_pzfx(paste(paper_path,
                                  "Figures for paper.xml",
                                  sep = ""),
                            table = table_name)
        ) %>%
        gather("condition", "cells") %>%
        mutate(description = table_name,
               met_potential = migration_tables$met_potential[i])

    if (data_type == "to_cytokine") {
        dataset <- dataset %>%
            filter(condition != "CXCL12" & condition != "OSM") %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("1% FBS", "Media + FBS") %>%
                       str_replace("CXCL8", "IL8")) %>%
            mutate(condition = factor(condition, levels = c("Media only",
                                                            "Media + FBS",
                                                            "IL6",
                                                            "IL8")))
    } else if (data_type == "to_inhibitors") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("Cond Media", "Cond. media (C.M.)") %>%
                       str_replace("CM \\+ sc144", "C.M. + SC144") %>%
                       str_replace("CM \\+ ladarixin", "C.M. + DF2156A") %>%
                       str_replace("CM \\+ both", "C.M. + both")) %>%
            mutate(condition = factor(condition, levels = c('Media only',
                                                            '1% FBS',
                                                            'Cond. media (C.M.)',
                                                            'C.M. + SC144',
                                                            'C.M. + DF2156A',
                                                            'C.M. + both')))
    } else if (data_type == "to_conditioned_media") {
        dataset <- dataset %>%
            mutate(condition =
                       str_replace(condition, "^Media$", "Media only") %>%
                       str_replace("1% FBS", "Media + FBS") %>%
                       str_replace("Cond Media", "Cond. media")) %>%
            mutate(condition = factor(condition, levels = c("Media only",
                                                            "Media + FBS",
                                                            "Cond. media")))
    }

    cmp_group <- if_else(data_type == "to_conditioned_media",
                         true = "Media . FBS",
                         false = "Media only")
    aov_test <- aov(cells ~ condition, data = dataset) %>%
        TukeyHSD(.)
    aov_test <- aov_test$condition %>%
        as.data.frame() %>%
        rownames_to_column("comparison") %>%
        as_tibble() %>%
        filter(grepl(cmp_group, comparison)) %>%
        mutate(comparison = str_remove(comparison, paste0("-*", cmp_group, "-*"))) %>%
        rename(p_adj = `p adj`) %>%
        filter(p_adj <= 0.05)

    obj_name <- paste0(cell_line, "_", data_type)
    assign(obj_name, dataset)
    saveRDS(obj_name, paste0("data/robj/", obj_name, ".rds"))

    plot_name <-
        ggplot(dataset, aes(condition, cells)) +
        geom_beeswarm(
            size = 2,
            color = "gray65",
            cex = 2
        ) +
        stat_summary(
            fun.data = mean_se,
            geom = "errorbar", 
            width = 0.3, 
            size = 1
        ) +
        stat_summary(
            fun = mean, 
            geom = "crossbar",
            width = 0.5,
            size = 0.5
        ) +
        labs(x = "",
             y = "Cells",
             title = cell_line) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

    if (nrow(aov_test) > 0) {
        plot_name <- plot_name +
            geom_text(data = aov_test,
                      aes(x = comparison,
                          y = max(dataset$cells) * 1.1,
                          label = paste0("p=", sprintf("%.2e", p_adj))),
                      size = 3)
    }

    ggsave(paste0("data/output/figures/",
                  cell_line,
                  "_",
                  data_type,
                  ".png"),
           width = 5,
           height = 5,
           plot = plot_name)
    print(plot_name)
}
```





# Conclusions

placeholder
 
