## Heterogeneity analysis

### Isolate the co-culture tumor and evaluate IL6 and CXCL8 expression

```{r}
# Load the processed culture data
cx <- qs::qread("output/seurat_objects/cx.qs")
tumor_cocx <- subset(cx, idents = "OS-CoCx")

# Remove the doublets that express epithelial markers
tumor_cocx <- subset(tumor_cocx, cells = readRDS("misc/doublets.rds"), invert = T)

tumor_cocx <- tumor_cocx %>%
    FindVariableFeatures() %>%
    RunPCA() %>%
    RunUMAP(
        dims = 1:20,
        n.neighbors = 20,
        min.dist = 0.1,
        local.connectivity = 5,
        spread = 1,
        verbose = FALSE) %>%
    FindNeighbors(verbose = FALSE) %>%
    FindClusters(verbose = FALSE, resolution = 0.25)

p1 <- r_dim_plot(tumor_cocx,
    title = "Tumor in Co-culture")
p2 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = IL6 > 3),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6 Expression >3") +
    theme(legend.position = "none") +
    coord_fixed()
p3 <- DimPlot(tumor_cocx,
        cells.highlight = WhichCells(tumor_cocx, expression = CXCL8 > 6),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CXCL8 Expression >6") +
    theme(legend.position = "none") +
    coord_fixed()
p4 <- r_feature_plot(tumor_cocx,
    "nCount_RNA",
    title = "Total Counts",
    max.cutoff = 12000)
p <- p1 | p2 | p3 | p4
p
stash_plot(p, "Hetero-1")

```


## Intratumoral Heterogeneity - Lung Met
```{r}
# pre_path <- "~/../lab/Counts_2/"
pre_path <- "C:/Users/rxr014/OneDrive - Nationwide Children's Hospital/BIScratch/"
post_path <- "/filtered_feature_bc_matrix"
samples <- tribble(
  ~id, ~model, ~mt_cutoff, ~count_cutoff, ~species,
  "S0281", "OS-17", 25, 160000, "^GRCh38-"
)

s <- 1
pdx_lungs <- tenx_load_qc(
    paste0(pre_path, samples$id[s], post_path),
    species_pattern = samples$species[s],
    violin_plot = FALSE
  ) %>%
  subset(
    percent.mt < samples$mt_cutoff[s] &
    nCount_RNA < samples$count_cutoff[s]
  ) %>%
  NormalizeData()
pdx_lungs$id <- samples$id[s]
pdx_lungs$model <- samples$model[s]

pdx_lungs <- pdx_lungs %>%
  FindVariableFeatures(nfeatures = 5000) %>%
  ScaleData() %>%
  RunPCA()

pdx_lungs <- pdx_lungs %>%
  FindNeighbors(reduction = "pca", k.param = 100L) %>%
  FindClusters(resolution = 0.8) %>%
  RunUMAP(reduction = "pca", dims = 1:30)

r_dim_plot(pdx_lungs)

pdx_lungs <- run_fdl(pdx_lungs)

r_dim_plot(pdx_lungs, reduction = "fdl") +
  labs(title = "OS-17 Lung Mets")

pdx_lungs <- kill_cc(pdx_lungs)

r_dim_plot(pdx_lungs,
    reduction = "fdl",
    group.by = "Phase",
    label = FALSE) +
  labs(title = "OS-17 Lung Mets - Cell Cycle") +
  theme(legend.position = "right")

Idents(pdx_lungs) <- pdx_lungs$seurat_clusters

r_feature_plot(pdx_lungs, "IL1R1", reduction = "fdl")
```

## Evaluate gene expression
```{r}
# Pull the transcription-factor regulated gene sets from msigdb
sigs <- msigdbr::msigdbr(species = "Homo sapiens",
    category = "H") %>%
  select(gs_name, gene_symbol)

degs <- subset(pdx_lungs, idents = c(0, 2, 5, 6)) %>%
  FindAllMarkers()

pathways <- future_lapply(unique(degs$cluster), function(i) {
  clust_degs <- degs %>%
    filter(cluster == i & p_val_adj < 0.05) %>%
    arrange(-avg_log2FC) %>%
    pull(avg_log2FC, name = gene)
  clusterProfiler::GSEA(clust_degs,
    TERM2GENE = sigs,
    nPermSimple = 10000) %>%
    mutate(p.adjust = -log10(p.adjust))
})
FeaturePlot()
plots <- lapply(seq_along(pathways), function(p) {
  enrichplot::dotplot(pathways[[p]] %>% filter(NES > 0),
    x = "NES",
    showCategory = 15,
    title = paste("Cluster", unique(degs$cluster)[p]))
})

patchwork::wrap_plots(plots, ncol = 2)

gs <- "HALLMARK_TNFA_SIGNALING_VIA_NFKB"

cell_ranks <- AUCell::AUCell_run(
  pdx_lungs[["RNA"]]$counts,
  sigs %>% filter(gs_name == gs) %>% pull(gene_symbol)
)
pdx_lungs$nfkb_activity <- AUCell::getAUC(cell_ranks)[1,]

FeaturePlot(pdx_lungs,
  "nfkb_activity",
  reduction = "fdl",
  cols = c("#EEEEEE", "#ae0600"),
  min.cutoff = 0.15,
  pt.size = 1,
  order = TRUE) +
  coord_fixed()

FeaturePlot(pdx_lungs,
  "IL1R1",
  reduction = "fdl",
  cols = c("#EEEEEE", "#ae0600"),
  min.cutoff = 1,
  pt.size = 1,
  order = TRUE) +
  coord_fixed()

```

## Test cell cycle regression
```{r}
pdx_lungs_cc <- pdx_lungs %>%
  ScaleData(
    vars.to.regress = c("S.Score", "G2M.Score"),
    features = rownames(x = pdx_lungs)) %>%
  RunPCA() %>%
  FindNeighbors(reduction = "pca", k.param = 100L) %>%
  FindClusters(resolution = 0.8) %>%
  RunUMAP(reduction = "pca", dims = 1:30)

r_dim_plot(pdx_lungs_cc, group.by = "Phase")

pdx_lungs_cc <- run_fdl(pdx_lungs_cc)

r_dim_plot(pdx_lungs_cc, group.by = "seurat_clusters", reduction = "fdl")
r_feature_plot(pdx_lungs_cc, "PAX7", reduction = "fdl")
```

## GSEA
```{r}
degs <- FindAllMarkers(pdx_lungs_cc) %>%
  filter(p_val_adj < 0.05) %>%
  arrange(-avg_log2FC)

pathways <- mclapply(unique(degs$cluster), function(clust) {
  genes <- degs %>%
    filter(cluster == clust)

  p <- clusterProfiler::gseGO(
    geneList = genes,
    OrgDb = org.Hs.eg.db::org.Hs.ed.db,
    ont = "BP",
    keyType = "SYMBOL",
    nPermSimple = 10000,
    eps = 0
  )

  return(p)
}, mc.cores = parallelly::availableCores())
```


## Intratumoral Heterogeneity - Primary Tumor
```{r}
samples <- tribble(
  ~id, ~model, ~mt_cutoff, ~count_cutoff, ~species,
  "S0037", "OS-17", 25, 160000, "^GRCh38-"
)

s <- 1
pdx_tibia <- tenx_load_qc(
    paste0(pre_path, samples$id[s], post_path),
    species_pattern = samples$species[s],
    violin_plot = FALSE
  ) %>%
  subset(
    percent.mt < samples$mt_cutoff[s] &
    nCount_RNA < samples$count_cutoff[s]
  ) %>%
  NormalizeData()
pdx_tibia$id <- samples$id[s]
pdx_tibia$model <- samples$model[s]

pdx_tibia <- pdx_tibia %>%
  FindVariableFeatures(nfeatures = 5000) %>%
  ScaleData() %>%
  RunPCA()

pdx_tibia <- pdx_tibia %>%
  FindNeighbors(reduction = "pca", k.param = 200L) %>%
  FindClusters(resolution = 0.8) %>%
  RunUMAP(reduction = "pca", dims = 1:30)

r_dim_plot(pdx_tibia)

pdx_tibia <- run_fdl(pdx_tibia)

r_dim_plot(pdx_tibia, reduction = "fdl") +
  labs(title = "OS-17 Primary Tibia")

pdx_tibia <- kill_cc(pdx_tibia)

r_dim_plot(pdx_tibia,
    reduction = "fdl",
    group.by = "Phase",
    label = FALSE) +
  labs(title = "OS-17 Primary Tibia - Cell Cycle") +
  theme(legend.position = "right")

Idents(pdx_tibia) <- pdx_tibia$seurat_clusters

r_feature_plot(pdx_tibia, "IL1R1", reduction = "fdl")
```

## Evaluate gene expression
```{r}
# Pull the transcription-factor regulated gene sets from msigdb
sigs <- msigdbr::msigdbr(species = "Homo sapiens",
    category = "H") %>%
  select(gs_name, gene_symbol)

# degs <- subset(pdx_tibia, idents = c(0, 2, 5, 6)) %>%
#   FindAllMarkers()

# pathways <- future_lapply(unique(degs$cluster), function(i) {
#   clust_degs <- degs %>%
#     filter(cluster == i & p_val_adj < 0.05) %>%
#     arrange(-avg_log2FC) %>%
#     pull(avg_log2FC, name = gene)
#   clusterProfiler::GSEA(clust_degs,
#     TERM2GENE = sigs,
#     nPermSimple = 10000) %>%
#     mutate(p.adjust = -log10(p.adjust))
# })

# plots <- lapply(seq_along(pathways), function(p) {
#   enrichplot::dotplot(pathways[[p]] %>% filter(NES > 0),
#     x = "NES",
#     showCategory = 15,
#     title = paste("Cluster", unique(degs$cluster)[p]))
# })

# patchwork::wrap_plots(plots, ncol = 2)

gs <- "HALLMARK_TNFA_SIGNALING_VIA_NFKB"

cell_ranks <- AUCell::AUCell_run(
  pdx_tibia[["RNA"]]$counts,
  sigs %>% filter(gs_name == gs) %>% pull(gene_symbol)
)
pdx_tibia$nfkb_activity <- AUCell::getAUC(cell_ranks)[1,]

FeaturePlot(pdx_tibia,
  "nfkb_activity",
  reduction = "fdl",
  cols = c("#EEEEEE", "#ae0600"),
  min.cutoff = 0.15,
  pt.size = 1,
  order = TRUE) +
  coord_fixed()

FeaturePlot(pdx_tibia,
  "IL1R1",
  reduction = "fdl",
  cols = c("#EEEEEE", "#ae0600"),
  min.cutoff = 1,
  pt.size = 1,
  order = TRUE) +
  coord_fixed()

```

## Test cell cycle regression
```{r}
pdx_tibia_cc <- pdx_tibia %>%
  ScaleData(
    vars.to.regress = c("S.Score", "G2M.Score"),
    features = rownames(x = pdx_tibia)) %>%
  RunPCA() %>%
  FindNeighbors(reduction = "pca", k.param = 100L) %>%
  FindClusters(resolution = 0.8) %>%
  RunUMAP(reduction = "pca", dims = 1:30)

r_dim_plot(pdx_tibia_cc, group.by = "Phase")

pdx_tibia_cc <- run_fdl(pdx_tibia_cc)

r_dim_plot(pdx_tibia_cc, group.by = "seurat_clusters", reduction = "fdl")
r_feature_plot(pdx_tibia_cc, "PAX7", reduction = "fdl")
```

## GSEA
```{r}
degs <- FindAllMarkers(pdx_tibia_cc) %>%
  filter(p_val_adj < 0.05) %>%
  arrange(-avg_log2FC)

pathways <- mclapply(unique(degs$cluster), function(clust) {
  genes <- degs %>%
    filter(cluster == clust)

  p <- clusterProfiler::gseGO(
    geneList = genes,
    OrgDb = org.Hs.eg.db::org.Hs.ed.db,
    ont = "BP",
    keyType = "SYMBOL",
    nPermSimple = 10000,
    eps = 0
  )

  return(p)
}, mc.cores = parallelly::availableCores())
```









### Load human met scRNA-seq and create Seurat objects

```{r}
deets_mets <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~src,
    "S0058", 50000, 14, "met-1",
    "S0059", 50000, 14, "met-2")

mets <- lapply(seq_len(nrow(deets_mets)), function(i) {
    s <- tenx_load_qc(paste0(scrna_path,
                deets_mets$id[i],
                "/filtered_feature_bc_matrix/"),
            violin_plot = FALSE) %>%
        subset(
            nCount_RNA < deets_mets$max_count[i] &
            percent.mt < deets_mets$max_percent_mt[i]) %>%
        process_seurat()
})

names(mets) <- deets_mets$id
for (i in seq_along(mets)) {
    mets[[i]]$src <- deets_mets$src[i]
}

tumor_cells <- readRDS("tumor_cells.rds")

mets <- merge(mets[[1]], mets[[2:length(mets)]],
    add.cell.ids = deets_mets$id) %>%
    process_seurat()
mets$tumor <- "stroma"
mets$tumor[names(mets$tumor) %in% readRDS("tumor_cells.rds")] <- "tumor"

r_dim_plot(mets, "Merged human lung metastases")
r_dim_plot(mets, group.by = "tumor", "Cells identified as tumor")

mets_tumor <- subset(mets, tumor == "tumor") %>%
    process_seurat()

p1 <- r_dim_plot(mets_tumor, "Isolated tumor cells")
p2 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = IL6 > 0),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "IL6") +
    theme(legend.position = "none") +
    coord_fixed()
p3 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = CXCL8 > 1),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CXCL8") +
    theme(legend.position = "none") +
    coord_fixed()
p4 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = MKI67 > 1),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "MKI67") +
    theme(legend.position = "none") +
    coord_fixed()
p5 <- DimPlot(mets_tumor,
        cells.highlight = WhichCells(mets_tumor, expression = CDKN1A > 1.5),
        cols.highlight = "#1F00B8",
        pt.size = 1) +
    labs(title = "CDKN1A") +
    theme(legend.position = "none") +
    coord_fixed()
p1 | p2 | p3 | p4 | p5
```

### Check for co-occurrence of IL6, CXCL8, p21, and Ki67 in human lung mets

```{r}
tmets <- list()
for (i in c("S0058", "S0059")) {
    r_dim_plot(mets[[i]], title = paste(i, "Clusters"))
    r_feature_plot(mets[[i]], "IL6", title = paste(i, "COL1A1 Expression"))
    DimPlot(mets[[i]],
            cells.highlight = WhichCells(mets[[i]], expression = IL6 > 1),
            cols.highlight = "#1F00B8",
            pt.size = 1) +
        labs(title = paste(i, "IL6 Expression >2")) +
        theme(legend.position = "none") +
        coord_fixed()

    # Isolate the tumor cells (list previously generated)
    # tmets[[i]] <- subset(mets[[i]],
    #     cells = readRDS(paste0(i, "TumorCells.rds")))

    # tmets[[i]] <- process_seurat(tmets[[i]])
    tmets[[i]] <- mets[[i]]
    r_dim_plot(tmets[[i]], title = paste(i, "Tumor Subclusters"))
    r_feature_plot(tmets[[i]], "IL6", title = paste(i, "IL6 Expression"))
    r_feature_plot(tmets[[i]], "CXCL8", title = paste(i, "CXCL8 Expression"))
    DimPlot(tmets[[i]],
            cells.highlight = WhichCells(tmets[[i]], expression = IL6 > 1),
            cols.highlight = "#1F00B8",
            pt.size = 1) +
        labs(title = paste(i, "IL6 Expression >1")) +
        theme(legend.position = "none") +
        coord_fixed()
}

tumor_cocx <- SCTransform(tumor_cocx)
s <- c("counts", "data", "scale.data")
p <- lapply(s, function(i) {
    t <- GetAssayData(tumor_cocx, slot = i) %>%
        as.data.frame() %>%
        t() %>%
        as.data.frame()

    return(
        ggplot(t, aes(IL6, CDKN1A)) +
            geom_point()
    )
})
p[[1]] | p[[2]] | p[[3]]

```
