## NicheNET analysis 
Starting with tumor (OS-17) and lung (HBEC) cells cultured either as monocultures or co-cultures, leverage the NicheNET algorithm to ask the question:

What ligands produced by epithelial cells
would bind receptors in osteosarcoma cells
to cause the target gene changes that are observed in the co-culture?

### Load and organize the co-culture data

```{r , eval=TRUE}
# scrna_path <- "../../../Counts_2/"
scrna_path <- "C:/Users/rxr014/OneDrive - Nationwide Children's Hospital/BIScratch/"

deets_cocx <- tribble(
    ~id, ~max_count, ~max_percent_mt, ~file,
    "S0001", 16000, 14, "S0001",
    "S0005", 16000, 14, "S0005")

cocx <- future_lapply(seq_len(nrow(deets_cocx)), function(i) {
    test <- tenx_load_qc(paste0(scrna_path,
                        deets_cocx$file[i],
                        "/filtered_feature_bc_matrix/"),
                 violin_plot = FALSE) %>%
        subset(nCount_RNA < deets_cocx$max_count[i] &
                   percent.mt < deets_cocx$max_percent_mt[i]) %>%
        process_seurat()
})

names(cocx) <- deets_cocx$id
for (i in seq_along(cocx)) {
    cocx[[i]]$src <- deets_cocx$id[i]
}
```

### Intercellular signaling analysis on co-culture data
Which signals produced by lung epithelial cells elicit the transcriptional changes that we see in osteosarcoma cells in co-culture?
In NicheNET, define the target gene set by DEG analysis comparing osteosarcoma cells in co-culture to cultured alone.
Define senders as the epithelial cells and receivers as the osteosarcoma cells.

```{r, eval=TRUE}
cx <- merge(
    cocx$S0001,
    y = cocx$S0005,
    add.cell.ids = c("monocx", "cocx"),
    project = "InterSignaling") %>%
    JoinLayers()

cx <- process_seurat(cx, resolution  = 0.4)

p1 <- r_dim_plot(cx,
    group.by = "src",
    title = "Source")
p2 <- r_dim_plot(cx, "Clusters")
p3 <- r_feature_plot(cx, features = "KRT19", title = "KRT19")
p4 <- r_feature_plot(cx, features = "COL1A1", title = "COL1A1")

p <- p1 | p2 | p3 | p4
p
stash_plot(p, "Co-culture")

# Suspect that clusters 7 and 8 are groups of doublets, remove
cx <- subset(cx, idents = c(7, 8), invert = T)

cx <- RenameIdents(cx,
    `5` = "OS-MonoCx",
    `1` = "OS-CoCx", `6` = "OS-CoCx",
    `0` = "HBEC-CoCx", `3` = "HBEC-CoCx", `4` = "HBEC-CoCx",
    `2` = "HBEC-MonoCx")

p <- r_dim_plot(cx, title = "OS-17 + HBEC, assignment")
p
stash_plot(p, "CoCx_assignments")

# Save off the processed Seurat object
qs::qsave(cx, "output/seurat_objects/cx.qs")

target_genes <- find_tar_genes(cx,
    id1 = "OS-CoCx",
    id2 = "OS-MonoCx",
    logfc = 2)

LRT_analysis_1 <- find_ligands(cx,
    target_genes,
    senders = c("HBEC-CoCx", "HBEC-MonoCx"),
    receiver = c("OS-CoCx", "OS-MonoCx"),
    rec_pct = 0.07,
    stringency = "strict")

p <- plot_complex_heatmap(LRT_analysis_1,
    grid_color_high = "#004466")
p
stash_plot(p, "Nichenet_heatmap")

p <- LRT_analysis_1[[3]] +
    scale_fill_gradient(low = "gray98", high = "#006600")
p
stash_plot(p, "Nichenet_targets")
```

### Evaluate changes in the production of ligands with co-culture

```{r}
ligands <- c(
    "TGFB1",
    "IL1A",
    "GMFB",
    "MMP9",
    "IL6",
    "CXCL8",
    "ADAM17",
    "HBEGF",
    "GAS6",
    "SAA1",
    "YARS",
    "PTHLH",
    "TGFA"
)

cx$assignment <- Idents(cx)
cx$assignment <- cx$assignment %>%
    factor(levels = c(
        "HBEC-MonoCx",
        "HBEC-CoCx",
        "OS-MonoCx",
        "OS-CoCx"
    ))

p <- DotPlot(cx, ligands, cols = "RdYlBu", group.by = "assignment") +
    coord_flip() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
p
stash_plot(p, "Co-cx_dotplot")
```
